{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2019-01-29 papers/methods/paper.tex
Elaborating on sec 4.5

The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{eqnarray}
m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{eqnarray}
What then remains is determining the functional form of $P(S|m_B)$. For the treatment of most surveys, we find that the error function, which smoothly transitions from some constant efficiency down to zero, is sufficient. Formally, this gives
\begin{equation}
P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
\end{equation}
where $\Phi^c$ is the complementary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function. The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}. However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a better analytic form, taking the form
\begin{equation}
P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
\end{equation}

@@

With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
\begin{eqnarray}
d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}
and is incorporated into our likelihood. \added{ Note that the above efficiencies utilise the common form of the normal distribution rather than the conditional probability notation found previously in this work.} This is illustrated in Figure \ref{fig:efficiency}. Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency. Further details on this choice are given in Section \ref{sec:simdes}.

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{efficiency.pdf}
	\end{center}
	\caption{The efficiency for supernova discovery at an arbitrary redshift. Shown in both panels in dashed blue is the SN~Ia population distribution, which takes the form of a normal distribution. The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), as functions of apparent magnitude. The survey efficiency, given the SN~Ia population, is shown as a solid line in both panels, and the probability of observing a SN~Ia is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated. This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.}
	\label{fig:efficiency}
\end{figure}

@@

Having laid out each individual aspect of the model, the relationships between variables and our treatment of uncertainty, here we summarise the relationships in our model likelihood mathematically. In this likelihood summary, we leave out sample selection for simplicity. Our likelihood is thus given as:

\begin{align}
&P({\rm data}|\theta) = \notag \\
&\int\, d\mu\, d\eta \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta) \notag \\
& P(\mu |z, \Omega_m, w) 
\end{align}
}

@@

As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{eqnarray}
m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
\sigma^*_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
\end{eqnarray}
Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.

>>>>

The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{eqnarray}
m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{eqnarray}
What then remains is determining the functional form of $P(S|m_B)$. For the treatment of most surveys, we find that the error function, which smoothly transitions from some constant efficiency down to zero, is sufficient. Formally, this gives
\begin{equation}
P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}), \label{eq:selcdf}
\end{equation}
where $\Phi^c$ is the complementary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function. The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}. However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a better analytic form, taking the form
\begin{equation}
P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}). \label{eq:selnorm}
\end{equation}

@@

With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
\begin{eqnarray}
d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-z}
\end{eqnarray}
and is incorporated into our likelihood. \added{ Note that the above efficiencies utilise the common form of the normal distribution rather than the conditional probability notation found previously in this work.} This is illustrated in Figure \ref{fig:efficiency}. Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency. Further details on this choice are given in Section \ref{sec:simdes}.

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{efficiency.pdf}
	\end{center}
	\caption{The efficiency for supernova discovery at an arbitrary redshift. Shown in both panels in dashed blue is the SN~Ia population distribution, which takes the form of a normal distribution. The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), as functions of apparent magnitude. The survey efficiency, given the SN~Ia population, is shown as a solid line in both panels, and the probability of observing a SN~Ia is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated. This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-z}.}
	\label{fig:efficiency}
\end{figure}

@@

Having laid out each individual aspect of the model, the relationships between variables and our treatment of uncertainty, here we summarise the relationships in our model mathematically. In this summary of $P({\rm data} | \theta)$, we leave out sample selection for simplicity. Referring to  equation \eqref{eq:main}, the relationships with $P({\rm data} | \theta)$ are as follows:

\begin{align}
&P({\rm data}|\theta) = \notag \\
&\int\, d\mu\, d\eta\, dz \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta) \notag \\
& P(\mu |z, \Omega_m, w) 
\end{align}

The denominator of equation \eqref{eq:main} is then given by either equation \eqref{eq:seldes} or \eqref{eq:sellow-z} depending on the survey, and similarly $P(S|{\rm data}, \theta)$ is given by equation \eqref{eq:selnorm} or \eqref{eq:selcdf} respectively. Combining all of these gives us our full model likelihood with selection effects accounted for.

}

@@

As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{eqnarray}
m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
\sigma^*_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
\end{eqnarray}
Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-z}.
